---
title       : Lobbyism in the U.S.  
subtitle    : Finanical Contributions in Election Years by the National Rifle Association
author      : B00H
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Motivation

Lobbying - trying to influence decision by the government, particularly legislators, plays a [major role in U.S. American politics](http://topics.nytimes.com/top/reference/timestopics/subjects/l/lobbying_and_lobbyists/index.html). One of the major players is the The National Rifle Association (NRA), an American nonprofit organization dedicated to protect the Second Amendment - the right to keep and bear arms. The NRA is opposed to any kind of gun regulation, including background checks, registration of fire arms, or restriction of assault weapon ownership, even though [the majority of US citizens supports stricter gun controls](http://www.gallup.com/poll/1645/guns.aspx). It exerts its political influence not only through ads (e.g., the most recent campaign against former New York mayor Michael Bloomberg and his gun control efforts), but also financial contributions to political campaigns. 

The [app](http://noname.shinyapps.io/DataProducts) aims to visualize the amount and the number of individual financial contributions by the NRA in election years 2000 though 2014. It provides some background information on the total number and amount (in US$) of individual contributions across US states. The interactive plots allow users to select U.S. states and to plot NRA contributions (both number and amount). 

The data was compiled and published by the [National Institute on Money in State Politics](http://www.followthemoney.org/), and the [data](https://github.com/B00H/DataProducts) was downloaded on 11 Sept 2014. 

---

## What's there to explore?

6 tabs: 

* __Introduction__: background information and motivation for app,

* Overview: plots total contributions (both number and amount) across U.S. states,

* US$ by state: plots total amount of contributions in millions US$ by state, whereby states are chosen by users (pre-selected: California),

* Contributions by state: plots total number of contributions in thousands by state, whereby states are chosen by users (pre-selected: California),

* Who gets the most?: plots number and amount of contributions across elections years by state,

* Code: link to code on [Github](https://github.com/B00H/DataProducts)

---

## Why should you explore the app? 

To get an idea how much money the National Rifle Organization has been spending in election years 2000 to 2014 across all and in individual U.S. states. 

To get an idea why Michael Bloomberg is investing a (significant amount of his private fortune)[http://www.nytimes.com/2014/04/16/us/bloomberg-plans-a-50-million-challenge-to-the-nra.html?_r=0] to support gun control measures.

To ponder why the National Rifle Association has been spending so much more money in California than other U.S. States. 

---

## Still need an incentive? 

```{r loadlibs, cache=TRUE, echo=FALSE, message=FALSE, results='hide'}
library(tidyr)
library(dplyr)
library(reshape2)
library(ggplot2)
## National Rifle Association donations between 2000 and 2014 by state. 
nra_raw <- read.delim("./FollowTheMoneyDownload20140911(030911).txt", 
                      stringsAsFactors=FALSE, colClasses = c("character"))
## Remove duplicate columns. 
nra_less <- nra_raw[, -c(1,2, 3, 5, 6)]
## Rename variables.
names(nra_less) <- c("election_state", "election_year", "contributions", "USD")
## Factorize, make numeric. 
nra_less$election_state <- as.factor(nra_less$election_state)
nra_less$election_year <- as.factor(as.numeric(nra_less$election_year))
nra_less$USD <- as.numeric(nra_less$USD)
nra_less$contributions <- as.numeric(nra_less$contributions)
## Order by year. 
nraFinal <- nra_less[order(nra_less$election_year), ]
## Remove 2015 entries, as zero. 
nraFinal <- nraFinal[-c(753:756),]
## Plot USD total by state. 
## Aggregate.
USD_State <- aggregate(nraFinal$USD, by = list(nraFinal$election_state), FUN = sum)
## Rename variables.
names(USD_State) <- c("state", "USD")
## Reorder levels. 
USD_State$state <- reorder(USD_State$state, USD_State$USD)
## Transform USD for better readability of y axis. 
USD_State$USD <- USD_State$USD/1000000
```

```{r plot, cache=TRUE, message=FALSE, echo=FALSE, fig.width=13, fig.height=7}
ggplot(USD_State, aes(x=state, y=USD)) + 
        geom_bar(stat="identity") +
        xlab("US State") +
        ylab("NRA donations in millions US$") +
        theme(axis.text.x = element_text(color = "black", size = 12, angle = 45), 
              axis.text.y = element_text(color = "black", size = 15),
              axis.title.x = element_text(color = "black", size = 18),
              axis.title.y = element_text(color = "black", size = 18, vjust = 1.5),
              panel.grid.minor.x = element_blank(), panel.grid.major.x=element_blank())
```

Go to the [app](http://noname.shinyapps.io/DataProducts).


